---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data

```{r}
library(shiny)
library(plotly)
library(dplyr)
library(ggplot2)
theme_set(theme_light())
library(readr)
library(lubridate)
library(tidytuesdayR)
library(forcats)
library(scales)

tt <- tt_load("2021-03-16")
tt
tt$games
```


```{r}
games <- tt$games %>%
  mutate(avg_peak_perc = parse_number(avg_peak_perc) / 100) %>%
  mutate(date = ymd(paste(year, month, 1))) %>%
  filter(date > min(date)) %>%
  mutate(month = fct_reorder(month, month(date))) %>%
  mutate(gamename = fct_reorder(gamename, -avg))
games
```

```{r}
game_names <- levels(games$gamename)

games %>%
  count(date) %>%
  ggplot(aes(date, n)) +
  geom_col()

```

```{r}
summarize_games <- function(tbl) {
  tbl %>%
    summarize(median_avg = median(avg),
              median_peak = median(peak))
}


games %>%
  group_by(date) %>%
  summarize_games() %>%
  ggplot(aes(date, median_avg)) +
  geom_line() +
  expand_limits(y = 0) +
  labs(x = "Month",
       y = "Median popularity of a game on Steam")

```

```{r}
games %>%
  filter(date < "2020-01-01") %>%
  group_by(month) %>%
  summarize_games() %>%
  ggplot(aes(month, median_avg)) +
  geom_line(group = 1) +
  expand_limits(y = 0) +
  labs(x = "Month of year",
       y = "Median popularity of a game on Steam",
       title = "Seasonal trend in Steam games",
       subtitle = "From 2012-2019")
```


```{r}
games %>%
  group_by(year, month) %>%
  summarize_games() %>%
  ggplot(aes(month, median_avg)) +
  geom_line(aes(color = factor(year), group = year)) +
  expand_limits(y = 0) +
  labs(x = "Month of year",
       y = "Median popularity of a game on Steam",
       title = "Seasonal trend in Steam games",
       color = "year")
```


```{r}
games %>%
  group_by(date) %>%
  summarize_games() %>%
  ggplot(aes(date, median_peak)) +
  geom_line() +
  expand_limits(y = 0)
```

Distribution of game popularity

```{r}
games %>%
  filter(date == max(date)) %>%
  ggplot(aes(avg)) +
  geom_histogram() +
  scale_x_log10(labels = comma,
                breaks = 10 ^ seq(0, 5)) +
  labs(x = "Average # of players across Feb 2021")
```


```{r}
games %>%
  filter(avg >= 100) %>%
  filter(date == max(date)) %>%
  ggplot(aes(avg_peak_perc)) +
  geom_histogram()

games %>%
  filter(avg >= 100) %>%
  filter(date == max(date)) %>%
  arrange((avg_peak_perc))

games %>%
  filter(avg >= 1000) %>%
  filter(date == "2021-02-01") %>%
  arrange((avg_peak_perc)) %>%
  ggplot(aes(avg, 1/ avg_peak_perc, label = gamename)) +
  geom_point() +
  geom_text(vjust = 1, hjust = 1, check_overlap = TRUE) +
  scale_y_log10() +
  scale_x_log10(labels = comma) +
  labs(x = "Average # of players in Feb 2021",
       y = "Ratio of Peak / Average")


games %>%
  filter(avg >= 1000,
         date == max(date)) %>%
  arrange((avg_peak_perc)) %>%
  ggplot(aes(avg, 1 / avg_peak_perc, label = gamename)) +
  geom_point() +
  geom_text(vjust = 1, hjust = 1, check_overlap = TRUE) +
  scale_y_log10() +
  scale_x_log10(labels = comma) +
  labs(x = "Average # of players in Feb 2021",
       y = "Ratio of Peak / Average")

# library(plotly)
# ggplotly(g)
games %>%
  filter(date == max(date)) %>%
  arrange(desc(avg))

```


```{r}
games %>%
  filter(fct_lump(gamename, 16, w = avg) != "Other") %>%
  mutate(gamename = fct_reorder(gamename, -avg)) %>%
  ggplot(aes(date, avg)) +
  geom_line() +
  expand_limits(y = 0) +
  scale_y_continuous(labels = comma) +
  facet_wrap(~ gamename, scales = "free_y") +
  labs(x = "Month",
       y = "Average players of this game in this month")
```


* Some games build popularity (Rust or Warframe or Rainbow Six Siege)
* Some games start big and drop off after (Grand Theft Auto V)
* Some games have sudden spikes that drop off (Terraria)
* Others have spikes that stick around a bit (PAYDAY 2)
* Some predate our 2012 data (Counter-Strike, Civilization)

# What became popular in the pandemic specifically?

```{r}
apr_feb_ratios <- games %>%
  filter(year == 2020) %>%
  select(gamename, month, avg, peak) %>%
  pivot_wider(names_from = month, values_from = c(avg, peak)) %>%
  select(gamename, contains("January"), contains("April"), contains("February")) %>%
  mutate(apr_feb_ratio = avg_April / avg_February) %>%
  filter(avg_January >= 100,
         avg_February >= 100) %>%
  arrange(desc(apr_feb_ratio))

apr_feb_ratios
```

```{r}
games %>%
  filter(date >= "2018-01-01") %>%
  inner_join(apr_feb_ratios %>%
               top_n(12, apr_feb_ratio), by = "gamename") %>%
  mutate(gamename = as.character(gamename)) %>%
  complete(gamename, date, fill = list(avg = 0)) %>%
  mutate(gamename = fct_reorder(gamename, -avg)) %>%
  ggplot(aes(date, avg)) +
  geom_line() +
  expand_limits(y = 0) +
  geom_vline(xintercept = as.Date("2020-03-01"),
             color = "red", lty = 2) +
  scale_y_continuous(labels = comma) +
  facet_wrap(~ gamename, scales = "free_y") +
  labs(x = "Month",
       y = "Average players of this game in this month")
```

```{r}
games %>% 
  count(gamename, wt = avg, sort = TRUE) %>% 
  pull(gamename) 
```

